Impacts on population indices if scientific surveys are excluded from marine protected areas

### 2023-08-18 text/main.tex
Update ms

% \noindent
% Possible target journals: CJFAS, ICES JMS, Fisheries Research, Ecological Applications?

\clearpage
% \linenumbers

@@

\noindent

Keywords:
groundfish,
index of abundance,
MPA,
scientific survey,
spatiotemporal,
spatio-temporal,
survey effort reduction

@@

We derived model-based indices of relative abundance or biomass using a spatiotemporal modelling approach with spatial and spatiotemporal Gaussian Markov random fields describing latent spatial and spatiotemporal processes \citep[e.g.,][]{shelton2014, thorson2015a, anderson2022}.
Depending on the species and survey, these models were configured as either delta models \citep{aitchison1955} with a logit link for a binomial model of encounter probability  and log link for a gamma model of positive catch weight \citep{schnute2003} or as \citet{tweedie1984} observation models with a log link for catch weight.
We included an offset term \citep[][p.~206]{mccullagh1989} for log area swept for the trawl models and log hook count for the long-line models.
Following a common model configuration for index standardization \citep[e.g.,][]{thorson2015a, thorson2019a}, our models include an independent intercept each year, a fixed spatial random field, and independent spatiotemporal random fields.
We reduced model complexity as needed based on diagnostics to reflect an analyst aiming for adequate model convergence and parsimony (Table~\ref{tab:model-configs}, Supporting Material).
We fit our models with the sdmTMB \textsf{R} package \citep{anderson2022}, which implements the models through TMB \citep{kristensen2016} with input matrices generated through \textsf{R}-INLA \citep{lindgren2015}.
Additional modelling details are available in the Supporting Material.

@@

    \item Accuracy: We calculated the annual relative error (RE) as $(\hat{\theta}_t - \theta_t) / \theta_t$, where $\hat{\theta}_t$ represents the index estimate from the restricted model in year $t$ and $\theta_t$ represents the index estimate from the status quo survey. We first centered $\hat{\theta}_t$ and $\theta_t$ by dividing them by their respective geometric means across years to avoid including consistent bias (an index from a shrunk survey domain should be lower than a full-survey domain index) and because these surveys are considered relative indices of abundance/biomass with an estimated catchability parameter that is estimated in stock assessments. We then calculated the absolute RE. The median absolute RE (MARE) represents a measure of accuracy loss---the expected magnitude of relative error in a typical year.

@@

We examined three potential predictors of changes in precision, accuracy, and trend bias:
(1) the CV of the status quo index, (2) the biomass or abundance proportion of a species inside the MPA network, and (3) trends in proportion of a species inside vs.\ outside the MPA.
For the first two, we fit generalized linear models (GLMs) with gamma observation error and log links with log-transformed predictor variables to identify patterns.
This implies that a proportional change in the predictor corresponds to a proportional change in the response.
For any species with slight proportional decreases in precision, we set these values to small positive values (0.01) for the purposes of this GLM because the log link requires positive data.
For the third, we compared the slope of relative error per decade against slope of density proportion within the MPA per decade for the most impacted survey, SYN WCHG and fit a linear regression.
% We did not fit these models to evaluate statistical significance (noting that we ignore uncertainty in the predictor and response variables) but to characterize the relationships.
Because these are proposed MPAs, they should not have had any causal effect on historical biomass or abundance.

@@

We then fit spatiotemporal models (omitting spatiotemporal random effects for simplicity, consistency, and estimation speed) with the same observation family and tested whether the upper 95\% CI on the slope parameter representing the effect of year was less than zero.
We repeated the simulation and estimation 120 times to calculate power: the proportion of times that the estimation models correctly identified a negative population trend.

\subsection*{Application to an MPA network off the West Coast of Canada}

@@

We applied the above approach to a specific MPA network and set of surveys off the West Coast of Canada.
The MPA network planning area, the Northern Shelf Bioregion (NSB), is approximately 102,000~km$^2$ and the proposed MPA network encompasses about 30\% of this area (30,493~km$^2$).
% TODO: this still correct?
The network process has been co-led by 15 partner First Nations, the province of British Columbia, and the Canadian federal government (Parks Canada, Environment and Climate Change Canada, and Fisheries and Oceans Canada). 
% CR: Parks Canada, EEEC and DFO? Might be worth mentioning, just because they represent a diversity of spaces (not just fisheries management).
% TODO: ER: is this correct?
Following a systematic conservation planning approach \citep{margules2000}, its design was guided by national and regional goals \citep{canada2011, canada2014} and centered on meeting network conservation objectives \citep{dfo2022networkactionplan}.
Of the six goals of the network, Goal 1 is of primary importance: ``protect and maintain marine biodiversity, ecological representation and special natural features.''
Additional goals and objectives include the protection of traditional use and cultural heritage, contributing to the protection of fisheries resources and their habitat, and providing opportunities for scientific research, education, and awareness \citep{canada2014}.
The proposed network is built off existing areas, where nearly 50\% of the draft scenario footprint already exists as some type of spatial protection (existing provincial or federal MPAs or fisheries closures in the form of Rockfish Conservation Areas) leaving $\sim$50\% of the network as new areas identified through the planning process \citep{dfo2022networkactionplan}.
The proposed NSB MPA network was announced by the planning partners in February 2023 with plans for iterative implementation; the first set of new MPAs are expected to be established by 2025.

The NSB MPA network proposes `Category 1' and `Category 2' zones.
Category 1 zones are proposed for implementation by 2025 and Category 2 zones by 2030 \citep{dfo2022networkactionplan}. 
In our analysis, we assume both zones are restricted from bottom trawl and long-line research surveys. 
From the existing Gwaii Haanas National Marine Conservation Area Reserve and Haida Heritage Site, we excluded trawl surveys from all ``Strict Protection'' zones, while only excluding the long-line survey from ``Strict Protection'' zones that overlap with Rockfish Conservation Areas. We do not exclude survey activity from remaining existing MPA zones where possible changes to activities are uncertain at this time \citep{dfo2022networkactionplan}.

@@

We analyzed data from four research surveys including the three synoptic bottom trawl surveys \citep[][Fig.~\ref{fig:map}]{sinclair2003syn, anderson2019synopsis}: Queen Charlotte Sound \citep[SYN QCS,][]{williams2018synqcs}, Hecate Strait \citep[SYN HS,][]{wyeth2018synhs}, and West Coast Haida Gwaii \citep[SYN WCHG,][]{williams2018synwchg}. We omitted two related surveys off the west coast of Vancouver Island because they are not affected by the proposed MPA network. The included surveys cover the continental shelf and upper slope off the northern part of the west coast of Canada (Fig.~\ref{fig:map}) and primarily target soft-bottom habitat. They have been conducted since 2003, are typically run in alternating years, and follow a random depth-stratified design with sampling units of 2 km $\times$ 2 km blocks with consistent gear and fishing protocols \citep{sinclair2003syn}. The unit of measurement for the synoptic trawl surveys is fish biomass per area swept.

@@

Full exclusion of scientific surveys from the proposed NSB MPA network zones would have resulted in a proportional loss of survey domain of \lostHBLL, \lostWCHG, \lostHS, and \lostQCS\ across the HBLL, SYN WCHG, SYN HS, and SYN QCS surveys, respectively (Fig.~\ref{fig:map}, Fig.~\ref{fig:strata}).
% If existing MPAs also excluded survey effort, \lostHBLLall, \lostWCHGall, and \lostQCSHSall\ of the survey domain would be excluded.
The impact of the proposed MPAs on indices was variable across species and surveys when looking at the time series (Fig.~\ref{fig:timeseries}, Fig.~\ref{fig:ts-hbll}--\ref{fig:ts-qcs}).
Overall trends were often similar between the status quo and shrunk scenarios, although the shrunk scenario had larger confidence intervals and increased interannual variability.
The increased interannual variability led to smoother generalized additive model fits in some cases---flattening some patterns that were apparent with the status quo indices (e.g.,~Fig.~\ref{fig:timeseries}).

@@

Restricting a survey nearly universally reduced precision (Fig.~\ref{fig:dotplot}a--d) and accuracy (Fig.~\ref{fig:dotplot}e--h) with respect to the status quo index and sometimes resulted in trend bias (Fig.~\ref{fig:dotplot}i--l).
Impacts on precision, accuracy, and trend bias were strongest in the SYN WCHG trawl survey (Fig.~\ref{fig:dotplot}a, e, i) followed by the HBLL OUT N survey (Fig.~\ref{fig:dotplot}b, f, j).
These two surveys contained species with the highest proportion of density within the proposed MPA network with proportions up to $\approx$0.5 (Fig.~\ref{fig:dotplot}).
For species with at least 15\% of their biomass or abundance density within the MPA, the average CV increase was \precisionWCHGshrunk\% in SYN WCHG and \precisionHBLLshrunk\% in HBLL OUT N (Fig.~\ref{fig:dotplot}a, b).
Accuracy loss, MARE---reflecting proportional error in a typical bad year---ranged up to \canaryrockfishSYNWCHGslopere\ (Canary Rockfish) in SYN WCHG and \southernrocksoleHBLLslopere\ (Southern Rock Sole) in HBLL OUT N with an average for species with at least 15\% density within the MPA of \accuracyWCHGshrunk\ and \accuracyHBLLshrunk\ in these surveys, respectively (Fig.~\ref{fig:dotplot}e, f).

Trends in RE ranged up to \canaryrockfishSYNWCHGslopere\ per decade for canary rockfish in SYN WCHG, meaning the index diverged from the status quo index by up to \canaryrockfishSYNWCHGsloperePerc\% per decade (Fig.~\ref{fig:dotplot}i).
In the HBLL OUT N, trends in RE per decade were largest for North Pacific Spiny Dogfish at \northpacificspinydogfishSYNWCHGslopere\ per decade.
Effects within the SYN QCS, where only four species had $>$15\% density within the MPA, were the smallest (Fig.~\ref{fig:dotplot}d, h, l).

@@

There were strong relationships between some covariates and our metrics of precision, accuracy, and trend bias for the shrunk-survey-domain indices (Figs.~\ref{fig:covariates} and ~\ref{fig:slopes}).
Indices with larger CVs tended to have greater losses of accuracy with survey restriction than indices starting with smaller CVs (Fig.~\ref{fig:covariates}a).
Across all surveys and species, this was at a rate of \covCvmaregeostat\ increase in MARE per 1\% increase in status quo index CV.
Species with a greater proportion of their stock within the MPAs suffered larger losses of accuracy and precision (Fig.~\ref{fig:covariates}b, c).
Across all surveys, a 1\% increase in proportion of a species within the MPA network resulted in an increase in MARE of \covMareGeo\ and an increase in CV of \covPrecisionGeo.

@@

Reductions in power to detect population declines were generally stronger in the SYN WCHG trawl survey than the HBLL OUT N longline survey (Fig.~\ref{fig:power}).
As examples, in the SYN WCHG trawl survey, the probability of detecting a 50\% decline for Redstripe Rockfish declined from 0.64 to 0.38.
Power to detect a 30\% decline for Redbanded Rockfish declined from 0.82 to 0.55.
% See 05-power.R 'NUMBERS FOR PAPER' to grab these hardcoded values
Reduction in power was generally 10 percentage points or smaller for the HBLL OUT N survey with the largest effects seen for Copper Rockfish and Tiger Rockfish.
Some species, such as Lingcod, saw minimal effects on power for both surveys.

@@

Our results quantified effects on precision and accuracy of survey indices.
Are these effects large enough to impact stock assessment?
In QCS and HS, the impacts of these survey exclusions are probably small enough that they are unlikely to have major consequences for stock assessment outcomes.
However, there may be impacts for some species within the SYN WCHG and the HBLL OUT N survey.
Our analysis found notable reductions in the power to detect 30\% and 50\% declines, which hint that these effects may meaningfully impact stock assessment estimates.
Ultimately, the impact on stock assessments will require further analysis---either with retrospective analyses degrading stock assessment input data, similar to that performed here, or through simulation.
A complete analysis would include closed-loop simulation \citep{smith1999, punt2016} under restricted survey scenarios to evaluate how these effects might impact the probability of achieving conservation and fishery objectives.

@@

We hypothesize that trend bias (trend in relative error) will have the greatest impact on stock assessment, recovery monitoring, and conservation agency assessment.
For some species, this trend bias was considerable.
For example, North Pacific Spiny Dogfish, Canary Rockfish, Redstripe Rockfish, and Petrale Sole had increases in relative error of $\approx$20--30\% per decade.
Index precision is sometimes ignored in simple stock assessments \citep{free2020} or adjusted through re-weighting schemes \citep{francis2011} and many index-based management procedures ignore precision in their calculations all together \citep{carruthers2016}.
A bias in trend, however, will directly impact index-based management procedures and affect the ``data'' to which stock assessments compare their calculations in the likelihood.
Additionally, if MPAs operate as hoped, the proportion of a species occurring inside the MPAs will initially increase over time, which would be expected to cause the restricted indices to underestimate the true overall species abundance resulting in a negative trend in relative error.
Therefore, trend bias may be even more present with real MPAs compared to our retrospective analysis where future MPAs had no causal mechanism to affect the past.
Fortunately, this potential for underestimation of abundance is consistent with the precautionary principle.

We evaluated impacts on indices of abundance if surveys are excluded from closed areas; however, survey exclusion would also impact biological samples such as age and length composition data.
The loss of such sampling within closed areas might be expected to result in larger impacts on stock assessment than degraded indices of abundance themselves (\citeauthor{yin2004} \citeyear{yin2004} but see \citeauthor{chen2003} \citeyear{chen2003}).
Composition data, and especially age composition data, are important inputs to modern integrated age structured stock assessments and provide high-quality information regarding recruitment, gear selectivity, growth, natural mortality, and relative depletion \citep{magnusson2007, ono2015}.
Furthermore, age and length samples are notoriously clustered \citep{hilbornwalters1992, francis2011}---fish of a similar age tend to be caught together---and removing spatial blocks from the survey domain could have a disproportionate impact on age composition data if the MPA happened to protect some non-random portion of the population (e.g., a nursery area for young fish, or a refuge or feeding area for older fish).
Finally, while spatiotemporal models are increasingly used for overall biomass or abundance density modelling (as done in this paper) and could use spatial information, covariates, or other surveys to alleviate the impact to indices, such approaches are less well-established for composition data \citep[but see][]{thorsonhaltuch2018}.

@@

There are several caveats to our analysis.
First, we performed a retrospective analysis on a historical dataset.
A simulation approach \citep[e.g.,][]{schnute2003, regular2020} could explore additional dimensions of the problem (e.g., systematically varying observation error, sampling intensity, or species distribution characteristics), derive potentially more generalizable results, and have a known truth for comparison instead of using the status quo index as a reference.
Conversely, using a historical dataset ``bakes'' in the ``messiness'' of reality (e.g., poorly understood outliers, clustering, movement, and spatially varying abundance trends) and makes the results pertinent to real stocks and real surveys where management decisions are made.
% TODO: next is kind of duplicated elsewhere?
Second, we suspect that surveys covering larger areas (e.g., Gulf of Alaska) might fare better under survey restriction when using the spatiotemporal modelling approach compared to the smaller surveys examined here (especially West Coast Haida Gwaii).
Compared to other surveys, West Coast Haida Gwaii spatiotemporal indices were more affected by survey exclusion than their design-based equivalent.
This might be because the survey is on the verge of lacking sufficient spatial data to support these spatiotemporal models.
Third, we found that indices were most impacted for species that had a higher relative abundance or biomass within the MPAs.
These species might also be expected to be most protected by the MPAs and it is possible this protection could offset some of the risk incurred by poorer survey coverage.
However, the degree of protection will depend on several factors including juvenile and adult species movement \citep[e.g.,][]{gerber2003, gruss2011} and larval dispersal distance \citep[e.g.,][]{botsford2003, planes2009} and knowledge about whether the protection is working will be challenging to establish without continued surveying within the MPAs.

@@

Should we use design- or model-based indices when faced with survey restrictions?
Overall, we found similar effects for design- and model-based indices.
However, we see several advantages of the spatiotemporal model approach in this context.
First, the models produced more precise indices in the status quo scenario \citep[sensu][]{thorson2015a}, and even though proportionally they were more impacted by restriction than the design-based index, they remained more precise, on average.
% TODO: some duplication with elsewhere here:
Second, spatiotemporal model indexes are likely to be less impacted for surveys covering larger spatial areas or with denser spatial coverage than some surveys considered here.
We suspect the spatiotemporal models were already struggling for WCHG where large spatial chunks were then removed.
If spatial information, covariate effects, or even spatially varying covariate effects \citep{hastie1993, barnett2021a, thorson2023svc} could be estimated well, we suspect a spatiotemporal model may be able to improve accuracy compared to omitting the MPA from the survey domain prediction grid by extrapolating into the closed area, but at the expense of precision.
% TODO: some duplication with the previous point here:
Third, the model-based approach affords the opportunity to leverage other data sources such as covariates \citep{johnson2019} and other surveys \citep[e.g.,][]{gruss2019, webster2020, monnahan2021} to enhance indices.
The approach provides a mechanism to make use of historical data in areas that later restrict surveys rather than discarding all historical data the does not match new stratum definitions.
Fourth, the model-based approach provides a statistical framework to make inference about the benefits of the MPAs through causal inference \citep[e.g.,][]{pearl2016}, assuming some sampling within the MPAs is maintained.

>>>>

\clearpage
\linenumbers

@@

\vspace{5mm}
\noindent
Keywords:
groundfish,
index of abundance,
MPA,
spatiotemporal or spatio-temporal,
survey effort reduction,
scientific survey exclusion

@@

We derived model-based indices of relative abundance or biomass using a geostatistical modelling approach with spatial and spatiotemporal Gaussian Markov random fields describing latent spatial and spatiotemporal processes \citep[e.g.,][]{shelton2014, thorson2015a, anderson2022}.
Depending on the species and survey, these models were configured as either delta models \citep{aitchison1955} with a logit link for a binomial model of encounter probability  and log link for a gamma model of positive catch weight \citep{schnute2003} or as \citet{tweedie1984} observation models with a log link for catch weight.
We included an offset term \citep[][p.~206]{mccullagh1989} for log area swept for the trawl models and log hook count for the long-line models.
Following a common model configuration for index standardization \citep[e.g.,][]{thorson2015a, thorson2019a}, our models included an independent intercept each year, a fixed spatial random field, and independent spatiotemporal random fields.
We reduced model complexity as needed based on diagnostics to reflect an analyst aiming for adequate model convergence and parsimony (Table~\ref{tab:model-configs}, Supporting Material).
We fit our models with the sdmTMB \textsf{R} package \citep{anderson2022}, which implements the models through TMB \citep{kristensen2016} with input matrices generated through \textsf{R}-INLA \citep{lindgren2015}.
Additional modelling details are available in the Supporting Material.

@@

    \item Accuracy: We calculated the annual relative error (RE) as $(\hat{\theta}_t - \theta_t) / \theta_t$, where $\hat{\theta}_t$ represents the index estimate from the restricted model in year $t$ and $\theta_t$ represents the index estimate from the status quo survey. We first centered $\hat{\theta}_t$ and $\theta_t$ by dividing them by their respective geometric means across years to avoid including consistent bias (an index from a shrunk survey domain should be lower than a full-survey domain index) and because surveys are often considered relative indices of abundance/biomass with an estimated catchability parameter that is estimated in stock assessments. We then calculated the absolute RE. The median absolute RE (MARE) represents a measure of accuracy loss---the expected magnitude of relative error in a typical year.

@@

We examined three potential predictors of changes in precision, accuracy, and trend bias:
(1) the CV of the status quo index, (2) the biomass or abundance proportion of a species inside the MPA network, and (3) trends in proportion of a species inside vs.\ outside the MPA.
For the first two, we fit generalized linear models (GLMs) with gamma observation error and log links with log-transformed predictor variables to identify patterns.
This implies that a proportional change in the predictor corresponds to a proportional change in the response.
For any species with slight proportional decreases in precision, we set these values to small positive values (0.01) for the purposes of this GLM because the log link requires positive data.
For the third, we compared the slope of relative error per decade against slope of density proportion within the MPA per decade for the most impacted survey(s) and fit a linear regression.
% We did not fit these models to evaluate statistical significance (noting that we ignore uncertainty in the predictor and response variables) but to characterize the relationships.
Because our analysis applies to proposed MPAs, the MPAs should not have had any causal effect on historical biomass or abundance.

@@

We then fit spatiotemporal models (omitting spatiotemporal random effects for simplicity, consistency, and estimation speed) with the same observation family and tested whether the upper 95\% confidence interval (CI) on the slope parameter representing the effect of year was less than zero.
We repeated the simulation and estimation 120 times to calculate power: the proportion of times that the estimation models correctly identified a negative population trend.

\subsection*{Application to an MPA network off the west coast of Canada}

@@

We applied the above approach to a specific MPA network and set of surveys off the west coast of Canada.
The MPA network planning area, the Northern Shelf Bioregion (NSB), is approximately 102,000~km$^2$ and the proposed MPA network encompasses about 30\% of this area (30,493~km$^2$).
% TODO: this still correct?
The network process has been co-led by 15 partner First Nations, the province of British Columbia, and the Canadian federal government (Parks Canada, Environment and Climate Change Canada, and Fisheries and Oceans Canada). 
% CR: Parks Canada, EEEC and DFO? Might be worth mentioning, just because they represent a diversity of spaces (not just fisheries management).
% TODO: ER: is this correct?
Following a systematic conservation planning approach \citep{margules2000}, its design was guided by national and regional goals \citep{canada2011, canada2014} and centered on meeting network conservation objectives \citep{dfo2022networkactionplan}.
Of the six goals of the network, Goal 1 is of primary importance: ``protect and maintain marine biodiversity, ecological representation and special natural features.''
Additional goals and objectives include the protection of traditional use and cultural heritage, contributing to the protection of fisheries resources and their habitat, and providing opportunities for scientific research, education, and awareness \citep{canada2014}.
The proposed network is built off existing areas, where nearly 50\% of the draft scenario footprint already exists as some type of spatial protection (existing provincial or federal MPAs or fisheries closures in the form of Rockfish Conservation Areas) leaving $\sim$50\% of the network as new areas identified through the planning process \citep{dfo2022networkactionplan}.
The proposed NSB MPA network was announced by the planning partners in February 2023 with plans for iterative implementation; the first set of new MPAs are expected to be established by 2025.

The NSB MPA network proposes `Category 1' and `Category 2' zones with the main distinction that Category 1 zones are proposed for implementation by 2025 and Category 2 zones by 2030 \citep{dfo2022networkactionplan}. 
In our analysis, we assume both zones are restricted from bottom trawl and long-line research surveys. 
From the existing Gwaii Haanas National Marine Conservation Area Reserve and Haida Heritage Site, we excluded trawl surveys from all ``Strict Protection'' zones, while only excluding the long-line survey from ``Strict Protection'' zones that overlap with Rockfish Conservation Areas. We do not exclude survey activity from remaining existing MPA zones where possible changes to activities are uncertain at this time \citep{dfo2022networkactionplan}.

@@

We analyzed data from four research surveys including the three synoptic bottom trawl surveys \citep[][Fig.~\ref{fig:map}]{sinclair2003syn, anderson2019synopsis}: 
% Queen Charlotte Sound \citep[SYN QCS,][]{williams2018synqcs}, 
% Hecate Strait \citep[SYN HS,][]{wyeth2018synhs}, and 
% West Coast Haida Gwaii \citep[SYN WCHG,][]{williams2018synwchg}. 
Queen Charlotte Sound (SYN QCS), 
Hecate Strait (SYN HS), and 
West Coast Haida Gwaii (SYN WCHG). 
% We omitted two related surveys off the west coast of Vancouver Island because they are not affected by the proposed MPA network. 
The included surveys cover the continental shelf and upper slope off the northern part of the west coast of Canada (Fig.~\ref{fig:map}) and primarily target soft-bottom habitat. They have been conducted since 2003, are typically run in alternating years, and follow a random depth-stratified design with sampling units of 2 km $\times$ 2 km blocks with consistent gear and fishing protocols \citep{sinclair2003syn}. The unit of measurement for the synoptic trawl surveys is fish biomass per area swept.

@@

Full exclusion of scientific surveys from the proposed NSB MPA network zones would have resulted in a proportional loss of survey domain of \lostHBLL, \lostWCHG, \lostHS, and \lostQCS\ across the HBLL OUT N, SYN WCHG, SYN HS, and SYN QCS surveys, respectively (Fig.~\ref{fig:map}, Fig.~\ref{fig:strata}).
% If existing MPAs also excluded survey effort, \lostHBLLall, \lostWCHGall, and \lostQCSHSall\ of the survey domain would be excluded.
The impact of the proposed MPAs on indices was variable across species and surveys when looking at the time series (Fig.~\ref{fig:timeseries}, Fig.~\ref{fig:ts-hbll}--\ref{fig:ts-qcs}).
Overall trends were often similar between the status quo and shrunk scenarios, although the shrunk scenario had larger confidence intervals and increased interannual variability.
The increased interannual variability led to smoother generalized additive model fits in some cases---flattening some patterns that were apparent with the status quo indices (e.g.,~Fig.~\ref{fig:timeseries}).

@@

Restricting a survey nearly universally reduced precision (Fig.~\ref{fig:dotplot}a--d) and accuracy (Fig.~\ref{fig:dotplot}e--h) with respect to the status quo index and sometimes resulted in trend bias (Fig.~\ref{fig:dotplot}i--l).
Impacts on precision, accuracy, and trend bias were strongest in the SYN WCHG trawl survey (Fig.~\ref{fig:dotplot}a,~e,~i) followed by the HBLL~OUT~N survey (Fig.~\ref{fig:dotplot}b,~f,~j).
These two surveys contained species with the highest proportion of density within the proposed MPA network with proportions up to $\approx$0.5 (Fig.~\ref{fig:dotplot}).
For species with at least 15\% of their biomass or abundance density within the MPA, the average CV increase was \precisionWCHGshrunk\% in SYN~WCHG and \precisionHBLLshrunk\% in HBLL~OUT~N (Fig.~\ref{fig:dotplot}a, b).
Accuracy loss, MARE---reflecting proportional error in a typical bad year---ranged up to \canaryrockfishSYNWCHGslopere\ (Canary Rockfish) in SYN WCHG and \southernrocksoleHBLLslopere\ (Southern Rock Sole) in HBLL~OUT~N with an average for species with at least 15\% density within the MPA of \accuracyWCHGshrunk\ and \accuracyHBLLshrunk\ in these surveys, respectively (Fig.~\ref{fig:dotplot}e,~f).

Trends in RE ranged up to \canaryrockfishSYNWCHGslopere\ per decade for canary rockfish in SYN WCHG, meaning the index diverged from the status quo index by up to \canaryrockfishSYNWCHGsloperePerc\% per decade (Fig.~\ref{fig:dotplot}i).
In the HBLL~OUT~N, trends in RE per decade were largest for North Pacific Spiny Dogfish at \northpacificspinydogfishSYNWCHGslopere\ per decade.
Effects within the SYN~QCS, where only four species had $>$15\% density within the MPA, were the smallest (Fig.~\ref{fig:dotplot}d,~h,~l).

@@

There were strong relationships between some covariates and our metrics of precision, accuracy, and trend bias for the shrunk-survey-domain indices (Figs.~\ref{fig:covariates} and ~\ref{fig:slopes}).
Indices with larger CVs tended to have greater losses of accuracy with survey restriction than indices starting with smaller CVs (Fig.~\ref{fig:covariates}a).
Across all surveys and species, this was at a rate of \covCvmaregeostat\ increase in MARE per 1\% increase in status quo index CV.
Species with a greater proportion of their density within the MPAs suffered larger losses of accuracy and precision (Fig.~\ref{fig:covariates}b, c).
Across all surveys, a 1\% increase in proportion of a species within the MPA network resulted in an increase in MARE of \covMareGeo\ and an increase in CV of \covPrecisionGeo.

@@

Reductions in power to detect population declines were generally stronger in the SYN WCHG trawl survey than the HBLL OUT N longline survey (Fig.~\ref{fig:power}).
As examples, in the SYN WCHG trawl survey, the probability of detecting a 50\% decline for Redstripe Rockfish based on a 95\%~CI declined from 0.64 to 0.38.
Power to detect a 30\% decline for Redbanded Rockfish declined from 0.82 to 0.55.
% See 05-power.R 'NUMBERS FOR PAPER' to grab these hardcoded values
Reduction in power was generally 10 percentage points or smaller for the HBLL OUT N survey with the largest effects seen for Copper Rockfish and Tiger Rockfish.
Some species, such as Lingcod, saw minimal effects on power for both surveys.

@@

Our results quantified effects on precision and accuracy of survey indices.
Are these effects large enough to impact stock assessment?
For SYN QCS and SYN HS, the impacts of these survey exclusions are probably small enough that they are unlikely to have major consequences for stock assessment outcomes.
However, there may be impacts for some species within the SYN WCHG and the HBLL OUT N survey.
Our analysis found notable reductions in the power to detect 30\% and 50\% declines, which hint that these effects may meaningfully impact stock assessment estimates.
Ultimately, the impact on stock assessments will require further analysis---either with retrospective analyses degrading stock assessment input data, similar to that performed here, or through simulation.
A complete analysis would include closed-loop simulation \citep{smith1999, punt2016} under restricted survey scenarios to evaluate how these effects might impact the probability of achieving conservation and fishery objectives.

@@

We hypothesize that trend bias (trend in relative error) will have the greatest impact on stock assessment, recovery monitoring, and conservation agency assessment.
For some species, this trend bias was considerable.
For example, North Pacific Spiny Dogfish, Canary Rockfish, Redstripe Rockfish, and Petrale Sole had increases in relative error of $\approx$20--30\% per decade.
Index precision is sometimes ignored in simple stock assessments \citep{free2020} or adjusted through re-weighting schemes \citep{francis2011} and many index-based management procedures ignore precision in their calculations all together \citep{carruthers2016}.
A bias in trend, however, will directly impact index-based management procedures and affect the ``data'' to which stock assessments compare their calculations in the likelihood.
Additionally, if MPAs operate as hoped, the proportion of a species occurring inside the MPAs will initially increase over time, which would be expected to cause the restricted indices to underestimate the true overall species abundance resulting in a negative trend in relative error.
Therefore, trend bias may be even more present with real MPAs compared to our retrospective analysis where future MPAs had no causal mechanism to affect the past.
Fortunately, this potential for underestimation of increases in abundance is consistent with the precautionary principle.

We evaluated impacts on indices of abundance if surveys are excluded from closed areas; however, survey exclusion would also impact biological samples such as age and length composition data.
The loss of such sampling within closed areas might be expected to result in larger impacts on stock assessment than degraded indices of abundance themselves (\citeauthor{yin2004} \citeyear{yin2004}, but see \citeauthor{chen2003} \citeyear{chen2003}).
Composition data, and especially age composition data, are important inputs to modern integrated age structured stock assessments and provide high-quality information regarding recruitment, gear selectivity, growth, natural mortality, and relative depletion \citep{magnusson2007, ono2015}.
Furthermore, age and length samples are notoriously clustered \citep{hilbornwalters1992, francis2011}---fish of a similar age tend to be caught together---and removing spatial blocks from the survey domain could have a disproportionate impact on age composition data if the MPA happened to protect some non-random portion of the population (e.g., a nursery area for young fish, or a refuge or feeding area for older fish).
Finally, while spatiotemporal models are increasingly used for overall biomass or abundance density modelling (as done in this paper) and could use spatial information, covariates, or other surveys to alleviate the impact to indices, such approaches are less well-established for composition data \citep[but see][]{thorsonhaltuch2018}.

@@

There are several caveats to our analysis.
First, we performed a retrospective analysis on a historical dataset.
A simulation approach \citep[e.g.,][]{schnute2003, regular2020} could explore additional dimensions of the problem (e.g., systematically varying observation error, sampling intensity, or species distribution characteristics), derive potentially more generalizable results, and have a known truth for comparison instead of using the status quo index as a reference.
Conversely, using a historical dataset ``bakes'' in the ``messiness'' of reality (e.g., poorly understood outliers, clustering, movement, and spatially varying abundance trends) and makes the results pertinent to real stocks and real surveys where management decisions are made.
% TODO: next is kind of duplicated elsewhere?
Second, we suspect that surveys covering larger areas (e.g., Gulf of Alaska) might fare better under survey restriction when using a spatiotemporal modelling approach compared to the smaller surveys examined here (especially West Coast Haida Gwaii).
Compared to other surveys, West Coast Haida Gwaii spatiotemporal indices were more affected by survey exclusion than their design-based equivalent.
This might be because the survey is on the verge of lacking sufficient spatial data to support these spatiotemporal models.
Third, we found that indices were most impacted for species that had a higher relative abundance or biomass within the MPAs.
These species might also be expected to be most protected by the MPAs and it is possible this protection could offset some of the risk incurred by poorer survey coverage.
However, the degree of protection will depend on several factors including juvenile and adult species movement \citep[e.g.,][]{gerber2003, gruss2011} and larval dispersal distance \citep[e.g.,][]{botsford2003, planes2009} and knowledge about whether the protection is working will be challenging to establish without continued surveying within the MPAs.

@@

Should we use design- or model-based indices when faced with survey restrictions?
Overall, we found similar effects for design- and model-based indices.
However, we see several advantages of the spatiotemporal model approach in this context.
First, the models produced more precise indices in the status quo scenario \citep[sensu][]{thorson2015a}, and even though proportionally they were more impacted by restriction than the design-based index, they remained more precise, on average.
% TODO: some duplication with elsewhere here:
Second, spatiotemporal model indexes are likely to be less impacted for surveys covering larger spatial areas or with denser spatial coverage than some surveys considered here.
We suspect the spatiotemporal models were already struggling for SYN WCHG where large spatial chunks were then removed.
If spatial information, covariate effects, or even spatially varying covariate effects \citep{hastie1993, barnett2021a, thorson2023svc} could be estimated well, we suspect a spatiotemporal model may be able to improve accuracy compared to omitting the MPA from the survey domain prediction grid by extrapolating into the closed area, but at the expense of precision.
% TODO: some duplication with the previous point here:
Third, the model-based approach affords the opportunity to leverage other data sources such as covariates \citep{johnson2019} and other surveys \citep[e.g.,][]{gruss2019, webster2020, monnahan2021} to enhance indices.
The approach provides a mechanism to make use of historical data in areas that later restrict surveys rather than discarding all historical data the does not match new stratum definitions.
Fourth, the model-based approach provides a statistical framework to make inference about the benefits of the MPAs through causal inference \citep[e.g.,][]{pearl2016}, assuming some sampling within the MPAs is maintained.
